---
title: "assignment-6"
author: "Gisela Noguera Toffoletti"
output: github_document
---

Assignment 6: Data import and tidy data
================

```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(janitor)
library(ggplot2)
```

<br>

## Exercise 1. Tibble and Data Import

Import the data frames listed below into R and
[parse](https://r4ds.had.co.nz/data-import.html#parsing-a-vector) the
columns appropriately when needed. Watch out for the formatting oddities
of each dataset. Print the results directly, **without** using
`kable()`.

**You only need to finish any three out of the five questions in this
exercise in order to get credit.**

<br>

#### 1.1 Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results. \[We didn’t have time to cover this in class, but look up how these functions work [here](https://r4ds.had.co.nz/tibbles.html#creating-tibbles)\]

```{r, echo=TRUE}
tribble(
  ~a, ~b, ~c,
  1,2.1, "apple",
  2, 3.2, "orange"
)

tibble(a = c(1,2),
       b = c(2.1,3.2),
       c = c("apple","orange"))
```


<br>
