---
title: "assignment-6"
author: "Gisela Noguera Toffoletti"
output: github_document
---

Assignment 6: Data import and tidy data
================

```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(janitor)
library(ggplot2)
```

<br>

## Exercise 1. Tibble and Data Import

Import the data frames listed below into R and
[parse](https://r4ds.had.co.nz/data-import.html#parsing-a-vector) the
columns appropriately when needed. Watch out for the formatting oddities
of each dataset. Print the results directly, **without** using
`kable()`.

**You only need to finish any three out of the five questions in this
exercise in order to get credit.**

<br>

#### 1.1 Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results. \[We didn’t have time to cover this in class, but look up how these functions work [here](https://r4ds.had.co.nz/tibbles.html#creating-tibbles)\]

```{r, echo=TRUE}
tribble(
  ~a, ~b, ~c,
  1,2.1, "apple",
  2, 3.2, "orange"
)

tibble(a = c(1,2),
       b = c(2.1,3.2),
       c = c("apple","orange"))
```


<br>

#### 1.2 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.

```{r, message=FALSE}
data_12 <- read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt", col_names = c("Name", "Weight", "Price"))
data_12
```

<br>

#### 1.3 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt` into R. Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.

```{r, warning=FALSE, message=FALSE}
data_13 <- read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt", skip = 2, na = c("?","Not Available"))
data_13 |> 
  mutate(`/Name/`= str_replace_all(`/Name/`,"/","")) |> 
  clean_names('upper_camel')
```


<br>

#### 1.4 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset4.txt` into R. Watch out for comments, units, and decimal marks (which are `,` in this case).

<br>

#### 1.5 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset5.txt` into R. Parse the columns properly. As a reminder, you can read about parsing date and time data [here](https://r4ds.had.co.nz/data-import.html#readr-datetimes). Write this imported and parsed data frame into a new csv file named `dataset5_new.csv` in your `problem_sets` folder.

<br>


