---
title: "assignment-6"
author: "Gisela Noguera Toffoletti"
output: github_document
---

Assignment 6: Data import and tidy data
================

```{r, warning = FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(janitor)
library(ggplot2)
```

<br>

## Exercise 1. Tibble and Data Import

Import the data frames listed below into R and
[parse](https://r4ds.had.co.nz/data-import.html#parsing-a-vector) the
columns appropriately when needed. Watch out for the formatting oddities
of each dataset. Print the results directly, **without** using
`kable()`.

**You only need to finish any three out of the five questions in this
exercise in order to get credit.**

<br>

#### 1.1 Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results. \[We didn’t have time to cover this in class, but look up how these functions work [here](https://r4ds.had.co.nz/tibbles.html#creating-tibbles)\]

```{r, echo=TRUE}
tribble(
  ~a, ~b, ~c,
  1,2.1, "apple",
  2, 3.2, "orange"
)

tibble(a = c(1,2),
       b = c(2.1,3.2),
       c = c("apple","orange"))
```


<br>

#### 1.2 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.

```{r, message=FALSE}
data_12 <- read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt", col_names = c("Name", "Weight", "Price"))
data_12
```

<br>

#### 1.3 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt` into R. Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.

```{r, warning=FALSE, message=FALSE}
data_13 <- read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt", skip = 2, na = c("?","Not Available"))
data_13 |> 
  mutate(`/Name/`= str_replace_all(`/Name/`,"/","")) |> 
  clean_names('upper_camel')
```


<br>

#### 1.4 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset4.txt` into R. Watch out for comments, units, and decimal marks (which are `,` in this case).

<br>

#### 1.5 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset5.txt` into R. Parse the columns properly. As a reminder, you can read about parsing date and time data [here](https://r4ds.had.co.nz/data-import.html#readr-datetimes). Write this imported and parsed data frame into a new csv file named `dataset5_new.csv` in your `problem_sets` folder.

<br>

## Exercise 2. Weather station

This dataset contains the weather and air quality data collected by a
weather station in Taiwan. It was obtained from the Environmental
Protection Administration, Executive Yuan, R.O.C. (Taiwan).

#### 2.1 Variable descriptions

- The text file
  `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station_notes.txt` contains descriptions of different variables collected by the station.
<br>
- Import it into R and print it in a table as shown below with `kable()`.

<br>
```{r, message=FALSE}
read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station_notes.txt", delim = "-") |> 
  kable()
```


<br>
`#` indicates invalid value by equipment inspection  
`*` indicates invalid value by program inspection  
`x` indicates invalid value by human inspection  
`NR` indicates no rainfall  
blank indicates no data

<br>

#### 2.2 Data tidying

- Import
  `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station.csv`
  into R. As you can see, this dataset is a classic example of untidy
  data: values of a variable (i.e. hour of the day) are stored as column
  names; variable names are stored in the `item` column.

- Clean this dataset up and restructure it into a tidy format.

- Parse the `date` variable into date format and parse `hour` into time.

- Turn all invalid values into `NA` and turn `NR` in rainfall into `0`.

- Parse all values into numbers.

- Show the first 6 rows and 10 columns of this cleaned dataset, as shown
  below, *without* using `kable()`.

*Hints: you don’t have to perform these tasks in the given order; also,
warning messages are not necessarily signs of trouble.*

<br>

Before cleaning:

```{r, message=FALSE, warning=FALSE}
data2 <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station.csv",  na = c("NA")) |> 
  mutate_all(as.character) |> 
  mutate(
    across(everything(), ~ replace(., . == "NR", 0)))

data2 |> 
  head()
#col_types = col(.defaults = "c") => import the data
#mutate(RAINFALL = ifelse(RAINFALL == "NR", 0, RAINFALL) |> 
#type_convert(col_types = cols(.default = col_double(), date = col_date(), station = col_character(), hour = col_time("%H")))
```

<br>

After cleaning:

```{r}
new_data <- data2 |> 
  pivot_longer(c(`00`:`23`), names_to = "hour") |> 
  pivot_wider(names_from = item, values_from = value) |> 
  mutate(
    hour = paste0(hour,":00"))
  
head(new_data)
```


<br>

#### 2.3 Using this cleaned dataset, plot the daily variation in ambient temperature on September 25, 2015, as shown below.

```{r}
new_data$hour <- factor(new_data$hour)
new_data |> 
  select(hour, AMB_TEMP, date) |> 
  filter(date == "2015-09-25") |> 
  ggplot(mapping = aes(x = hour, y = AMB_TEMP, group = 1)) +
  geom_line()
```


<br>

#### 2.4 Plot the daily average ambient temperature throughout the year with a **continuous line**, as shown below.

```{r}
new_data$date <- factor(new_data$date)
new_data |> 
  select(hour,AMB_TEMP,date) |> 
  ggplot(mapping = aes(x = date, y = AMB_TEMP, group = 1)) +
  geom_line() 
```


<br>

#### 2.5 Plot the total rainfall per month in a bar chart, as shown below.

*Hint: separating date into three columns might be helpful.*

```{r, warning=FALSE, message=FALSE}
new_data |> 
  mutate(
    month = format(as.Date(date), "%m"),
    RAINFALL = as.numeric(RAINFALL)) |> 
  group_by(month) |> 
  summarise(
    MonthlyRainfall = sum(RAINFALL, na.rm = T)
  ) |> 
  ggplot(mapping = aes(x = month, y = MonthlyRainfall)) +
  geom_col()
```

<br>

#### 2.6 Plot the per hour variation in PM2.5 in the first week of September with a **continuous line**, as shown below.

*Hint: uniting the date and hour and parsing the new variable might be
helpful.*

```{r}
new_data <- new_data |> 
  mutate(
    time = parse_datetime(paste(date, hour), format = "%Y-%m-%d %H:%M"))
new_data$time <- factor(new_data$time) 
new_data |> 
  filter(between(as.Date(time), as.Date("2015-01-01"), as.Date("2015-01-08"))) |> 
  ggplot(mapping = aes(x = time, y = PM2.5, group = 1)) +
  geom_line() 
```


<br>
